Parallel Information Set Generation for Kriegspiel

### 2012-06-27 sbac/paper.tex
Tables merge an, abstract, intro

\begin{abstract}
Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
consistent with a player's observations.  The ability to reason about the possible game history is critical to the
performance of game playing agents.  In this work, we address the parallelization of information set generation in the context
of Kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
engine and discuss its performance under different load balancing strategies and with different grain-size parameters.
We achieve speedups of over 500 on 1024 processors, far exceeding previous scalability results for game tree search applications.
\end{abstract}

@@

Information Set Generation (ISG) is an operation that game tree search agents
can use to identify nodes in the game tree that are consistent with their
observations~\cite{richards12reasoning}.  Because the operation is relatively
expensive and may be repeatedly applied by a decision-making agent, its
efficiency is a key factor in an agent's overall performance.  In this work, we
show that ISG for Kriegspiel can be efficiently parallelized.

@@

Information set generation is not the same problem as solving a game tree, although it is arguably a necessary
subroutine for quality game play in imperfect information games.  In this work, we describe an approach to generating
information sets for the game of Kriegspiel.

@@

\subsubsection{Iso-efficiency}
\begin{figure}[h]
\centering
\begin{minipage}{0.5\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.5\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
\end{minipage}
\vspace{-0.1in}
\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
(Bottom) Same data with a linear y-axis for the running time (to help see the Iso-efficiency.)}
\vspace{-0.2in}
\label{3depths}
\end{figure}

We next attempted to measure the Iso-efficiency [TODO:cite] of the problem---how is efficiency of the parallelization of the
algorithm affected by {\em size} of the problem?  To test this, we varied the number of plies in the test instance from
8 to 12. We used the work stealing load balancing strategy.  The results are shown in Figure~\ref{3depths}.   In the easiest case (depth 8), there isn't
enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
number of processors increases.

@@

\begin{table}[tphb]
\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & {\bf 177} & 178 & 184 & 493 \\
4 & 71 & 53 & {\bf 48} & 75 \\
16 & 34 & 19 & {\bf 12} & 16 \\
64 & 24 & 7 & {\bf 3.7} & 4.1 \\
256 & 24 & 4.4 & 1.2 & {\bf 1.0} \\
1024 & 26 & 3.8 & 0.64 & {\bf 0.40}
\end{tabular}
\caption{Running times, in seconds, to find full information set of size 764,209 at position 10 of Virgil-David 1969. Optimal grainsize (G) for each number of cores (PE) is shown in bold}
\vspace{-0.2in}
\label{tab:prob4}
\end{table}

\begin{table}[tphb]
\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & 617 & {\bf 612} & 650 & 3312 \\
4 & 213 & 170 & {\bf 158} & 367 \\
16 & 83 & 49 & {\bf 40} & 64 \\
64 & 54 & 19 & {\bf 11} & 14.9 \\
256 & 54 & 9.8 & 3.5 & {\bf 3.5} \\
1024 & 58 & 5.5 & 1.3 & {\bf 1.2} \\
\end{tabular}
\caption{Running times, in seconds, to find full information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006. Optimal grainsize for each number of cores is shown in bold.}
\vspace{-0.2in}
\label{tab:prob5}
\end{table}

\begin{table}[tphb]
\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & 1 & 0.99 & 0.96 & 0.36 \\ 
4 & 2.48 & 3.34 & 3.63 & 2.35 \\
16 & 5.19 & 9.27 & 13.8 & 10.7 \\
64 & 7.25 & 23.6 & 47.39 & 43.1 \\
256 & 7.1 & 40.1 & 145 & 172 \\
1024 & 6.61 & 46.5 & 276 & {\bf 443}
\end{tabular}
\caption{Speedups for Virgil-David 1969.}
\vspace{-0.2in}
\label{tab:prob4speedups}
\end{table}

\begin{table}[tphb]
\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & 1 & 1 & 0.95 & 0.19 \\
4 & 2.89 & 3.63 & 3.89 & 1.68 \\
16 & 7.41 & 12.4 & 15.1 & 9.57 \\
64 & 11.3 & 32.3 & 52.9 & 41.4 \\
256 & 11.3 & 62.9 & 172 & 174 \\
1024 & 10.6 & 111 & 464 & {\bf 536}
\end{tabular}
\caption{Speedups for Kbot-Darkboard 2006.}
\vspace{-0.2in}
\label{tab:prob5speedups}
\end{table}

>>>>

\begin{abstract}
Information Set Generation is the identification of the set of paths in an imperfect information game tree that are
consistent with a player's observations.  The ability to reason about the possible game history is critical to the
performance of game playing agents.  ISG represent a class of combinatorial search problems which is computationaly intensive but difficult to efficiently parallelize.

In this paper, we address the parallelization of information set generation in the context
of Kriegspiel (partially observable chess).  We implement the algorithm on top of a general purpose combinatorial search
engine and discuss its performance under different load balancing strategies and with different grain-size parameters.
We achieve speedups of over 500 on 1024 processors, far exceeding previous scalability results for game tree search applications.
\end{abstract}

@@

Information Set Generation (ISG) is an operation that game tree search agents
can use to identify nodes in the game tree that are consistent with their
observations~\cite{richards12reasoning}.  Because the operation is relatively
expensive and may be repeatedly applied by a decision-making agent, its
efficiency is a key factor in an agent's overall performance. Information set generation is not the same problem as solving a game tree, although it is arguably a necessary subroutine for quality game play in imperfect information games.

@@

In this work, we show that ISG for Kriegspiel can be efficiently parallelized. The primary contributions of this work are the following. We describe an approach to efficient parallelization of an application which is not typical of high performance computing domain. We present scalability results for our parallel implementation of information set generation algorithm for the game of Kriegspiel using real game problem instances. Further, we demonstrate the effect of load balancing and grain size on parallel performance of our solution.

@@

\subsubsection{Effect of Problem Size}
\begin{figure}[h]
\centering
\begin{minipage}{0.5\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.5\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
\end{minipage}
\vspace{-0.1in}
\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
(Bottom) Same data with a linear y-axis for the running time (to help see the Iso-efficiency.)}
\vspace{-0.2in}
\label{3depths}
\end{figure}

We next study the variation in the efficiency of the parallelization of the
algorithm with {\em size} of the problem?  To test this, we varied the number of plies in the test instance from
8 to 12. We used the work stealing load balancing strategy.  The results are shown in Figure~\ref{3depths}.   In the easiest case (depth 8), there isn't
enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
number of processors increases.

@@

\begin{table}[tphb]
\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & {\bf 177$|$1} & 178$|$0.99 & 1844$|$0.96 & 493$|$0.36 \\
4 & 71$|$2.48 & 53$|$3.34 & {\bf 48$|$3.63} & 75$|$2.35 \\
16 & 34$|$5.19 & 19 $|$9.27& {\bf 12$|$13.8} & 16$|$10.7 \\
64 & 24 $|$7.25 & 7 $|$23.6& {\bf 3.7$|$47.39} & 4.1$|$43.1 \\
256 & 24 $|$ 7.1& 4.4 $|$40.1 & 1.2 $|$145& {\bf 1.0$|$172} \\
1024 & 26 $|$6.61 & 3.8 $|$46.5& 0.64 $|$276& {\bf 0.40$|$443}
\end{tabular}
\caption{Running times in seconds and speedups (after $|$), to find full information set of size 764,209 at position 10 of Virgil-David 1969. Optimal grainsize (G) for each number of cores (PE) is shown in bold}
\vspace{-0.2in}
\label{tab:prob4}
\end{table}

\begin{table}[tphb]
\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & 617$|$1 & {\bf 612$|$1} & 650$|$0.95 & 3312$|$0.19 \\
4 & 213$|$2.89 & 170$|$3.63 & {\bf 158$|$3.89} & 367$|$1.68 \\
16 & 83$|$7.41 & 49$|$12.4 & {\bf 40$|$15.1} & 64$|$9.57 \\
64 & 54$|$11.3 & 19$|$32.3 & {\bf 11$|$52.9} & 14.9$|$41.4 \\
256 & 54$|$11.3 & 9.8$|$62.9 & 3.5$|$172 & {\bf 3.5$|$174} \\
1024 & 58$|$10.6 & 5.5$|$111 & 1.3$|$464 & {\bf 1.2$|$536} \\
\end{tabular}
\caption{Running times, in seconds and speedups (after $|$), to find full information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006. Optimal grainsize for each number of cores is shown in bold.}
\vspace{-0.2in}
\label{tab:prob5}
\end{table}
